Opening up the court (surface) in tennis grand slams

### 2020-08-02 paper_results/EDA/graphics_eda.Rmd
kayla update figure 5 to try new things with height vs aces

>>>>

#### Using ace percentage.... one dot per tournament (colored)

maxish <- function(x){
  xx <- na.omit(unique(x))
  return(ifelse(length(xx) > 1, max(xx), xx))
}

atp_servers <- c("Ivo Karlovic", "John Isner", "Rafael Nadal",
                 "Roger Federer", "Andy Murray")

wta_servers <- c("Karolina Pliskova", "Caroline Wozniacki",
                 "Venus Williams", "Serena Williams", "Madison Keys")

atp_ace_ht_tour <-   gs_players %>% 
  filter(league == "ATP") %>%
  dplyr::group_by(name, tournament) %>%
  dplyr::summarise(ht = maxish(ht),
                   avg_ace_perc = mean(ace/svpt, na.rm=TRUE),
                   med_ace_perc = median(ace/svpt, na.rm=TRUE),
                   count_n = n())

wta_ace_ht_tour <-   gs_players %>% 
  filter(league == "WTA") %>%
  dplyr::group_by(name, tournament) %>%
  dplyr::summarise(ht = maxish(ht),
                   avg_ace_perc = mean(ace/svpt, na.rm=TRUE),
                   med_ace_perc = median(ace/svpt, na.rm=TRUE),
                   count_n = n())

# atp_servers <- sapply(ungroup(atp_ace_ht_tour),
#                       function(x) return(sample(which(ungroup(atp_ace_ht_tour)$name == x), 2))) %>%
#   as.numeric() %>% ungroup(atp_ace_ht_tour)[., ]
# 
# wta_servers <- sapply(wta_ace_ht_tour,
#                       function(x) return(sample(which(wta_ace_ht_tour$name == x), 2))) %>%
#   as.numeric() %>% gs_players[., ]

p_atp_percent <- 
atp_ace_ht_tour %>% # could also try median
ggplot(aes(ht, med_ace_perc)) + 
  geom_jitter(aes(color = tournament), size=1.5) + 
  # scale_color_manual(" ", values = tournament_colors) +
  geom_smooth(aes(color = tournament), method = "lm") + 
  geom_label_repel(data=dplyr::filter(atp_ace_ht_tour, name %in% atp_servers),
            aes(ht, med_ace_perc, label=name, color = tournament),
            alpha=.8, seed=1234, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) + 
  geom_label_repel(data=dplyr::filter(atp_ace_ht_tour, name %in% atp_servers),
            aes(ht, med_ace_perc, label = name, color = tournament),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) +
  labs(x = "Height in cm", y = "Median ace% per match", title = "ATP") +
  my_theme +
  scale_color_manual("Tournament",
                                values = tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)\n",
                                           "French Open\n(Clay)\n",
                                           "US Open\n(DecoTurf)\n",
                                           "Wimbledon\n(Grass)\n")) +
# +
#   guides(color = guide_legend(override.aes = list(size = 7, shape=15))) + 
facet_wrap(~tournament) +
theme(legend.position="none")

p_atp_percent

if(save_graph){
ggsave("../plots/ace_ht_facet_atp.jpg", height=8, width=graphic_width_short)
}

p_wta_percent <- 
wta_ace_ht_tour %>% # could also try median
ggplot(aes(ht, med_ace_perc)) + 
  geom_jitter(aes(color = tournament), size=1.5) + 
  # scale_color_manual(" ", values = tournament_colors) +
  geom_smooth(aes(color = tournament), method = "lm") + 
  geom_label_repel(data=dplyr::filter(wta_ace_ht_tour, name %in% wta_servers),
            aes(ht, med_ace_perc, label=name, color = tournament),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = rep(c(-.1, .06, 0.06, 0.08, -0.1), 4),
            nudge_x = rep(c(-5, 0, 0, -7, 0), 4)) +
  geom_label_repel(data=dplyr::filter(wta_ace_ht_tour, name %in% wta_servers),
            aes(ht, med_ace_perc, label = name, color = tournament),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = rep(c(-.1, .06, 0.06, 0.08, -0.1), 4),
            nudge_x = rep(c(-5, 0, 0, -7, 0), 4)) +
  labs(x = "Height in cm", y = "Median ace% per match", title = "WTA") +
  my_theme +
  scale_color_manual("Tournament",
                                values = tournament_colors,
                                labels = c("Australian Open\n(Plexicushion)\n",
                                           "French Open\n(Clay)\n",
                                           "US Open\n(DecoTurf)\n",
                                           "Wimbledon\n(Grass)\n")) +
# +
#   guides(color = guide_legend(override.aes = list(size = 7, shape=15))) + 
facet_wrap(~tournament) +
theme(legend.position="none") + 
  ylim(c(-0.02, 0.2))

p_wta_percent

if(save_graph){
ggsave("../plots/ace_ht_facet_wta.jpg", height=8, width=graphic_width_short)
}

<!-- #### Using ace percentage.... one dot per tournament (black white) -->

<!-- ```{r, fig.width=16, fig.height=8, fig.align='center'} -->
<!-- set.seed(8) -->

<!-- maxish <- function(x){ -->
<!--   xx <- na.omit(unique(x)) -->
<!--   return(ifelse(length(xx) > 1, max(xx), xx)) -->
<!-- } -->

<!-- atp_servers <- c("Ivo Karlovic", "John Isner", "Rafael Nadal", -->
<!--                  "Roger Federer", "Andy Murray") -->

<!-- wta_servers <- c("Karolina Pliskova", "Caroline Wozniacki", -->
<!--                  "Venus Williams", "Serena Williams", "Madison Keys") -->

<!-- atp_ace_ht_tour <-   gs_players %>%  -->
<!--   filter(league == "ATP") %>% -->
<!--   dplyr::group_by(name, tournament) %>% -->
<!--   dplyr::summarise(ht = maxish(ht), -->
<!--                    avg_ace_perc = mean(ace/svpt, na.rm=TRUE), -->
<!--                    med_ace_perc = median(ace/svpt, na.rm=TRUE), -->
<!--                    count_n = n()) -->

<!-- wta_ace_ht_tour <-   gs_players %>%  -->
<!--   filter(league == "WTA") %>% -->
<!--   dplyr::group_by(name, tournament) %>% -->
<!--   dplyr::summarise(ht = maxish(ht), -->
<!--                    avg_ace_perc = mean(ace/svpt, na.rm=TRUE), -->
<!--                    med_ace_perc = median(ace/svpt, na.rm=TRUE), -->
<!--                    count_n = n()) -->

<!-- atp_servers <- sapply(atp_servers, -->
<!--                       function(x) return(sample(which(gs_players$name == x), 2))) %>% -->
<!--   as.numeric() %>%  -->
<!--   gs_players[., ]  %>% -->
<!--   dplyr::mutate(med_ace_perc = median(ace/svpt, na.rm=TRUE)) -->

<!-- wta_servers <- sapply(wta_servers, -->
<!--                       function(x) return(sample(which(gs_players$name == x), 2))) %>% -->
<!--   as.numeric() %>% gs_players[., ] %>% -->
<!--   dplyr::mutate(med_ace_perc = median(ace/svpt, na.rm=TRUE)) -->

<!-- p_atp_percent <-  -->
<!-- atp_ace_ht_tour %>% # could also try median -->
<!-- ggplot(aes(ht, med_ace_perc)) +  -->
<!--   geom_jitter( size=1.5) +  -->
<!--   # scale_color_manual(" ", values = tournament_colors) + -->
<!--   geom_smooth( method = "lm", color = "black") +  -->
<!--   geom_label_repel(data=atp_servers, -->
<!--             aes(ht, med_ace_perc, label=name), -->
<!--             alpha=.8, seed=1234, show.legend=FALSE, size = 7, -->
<!--             nudge_y = c(-.075, 0, 0, -0.05,0.02 ), -->
<!--             nudge_x = c(5, 5, -5, -3, 0)) +  -->
<!--   geom_label_repel(data=atp_servers, -->
<!--             aes(ht, med_ace_perc, label = name), -->
<!--             seed=1234, fill=NA, show.legend=FALSE, size = 7, -->
<!--             nudge_y = c(-.075, 0, 0, -0.05,0.02 ), -->
<!--             nudge_x = c(5, 5, -5, -3, 0)) + -->
<!--   labs(x = "Height in cm", y = "Median ace% per match", title = "ATP") + -->
<!--    my_theme # + -->
<!-- #   scale_color_manual("Tournament", -->
<!-- #                                 values = tournament_colors, -->
<!-- #                                 labels = c("Australian Open\n(Plexicushion)\n", -->
<!-- #                                            "French Open\n(Clay)\n", -->
<!-- #                                            "US Open\n(DecoTurf)\n", -->
<!-- #                                            "Wimbledon\n(Grass)\n")) + -->
<!-- # # + -->
<!-- # #   guides(color = guide_legend(override.aes = list(size = 7, shape=15))) +  -->
<!-- # facet_wrap(~tournament) + -->
<!-- # theme(legend.position="none") -->

<!-- p_atp_percent -->

<!-- p_wta_percent <-  -->
<!-- wta_ace_ht_tour %>% # could also try median -->
<!-- ggplot(aes(ht, med_ace_perc)) +  -->
<!--   geom_jitter(size=1.5) +  -->
<!--   # scale_color_manual(" ", values = tournament_colors) + -->
<!--   geom_smooth(method = "lm", color = "black") +  -->
<!--   geom_label_repel(data=wta_servers, -->
<!--             aes(ht, med_ace_perc, label=name), -->
<!--             seed=1234, fill=NA, show.legend=FALSE, size = 5) + -->
<!--             # nudge_y = rep(c(-.1, .06, 0.06, 0.08, -0.1), 4), -->
<!--             # nudge_x = rep(c(-5, 0, 0, -7, 0), 4)) + -->
<!--   geom_label_repel(data=wta_servers, -->
<!--             aes(ht, med_ace_perc, label = name), -->
<!--             seed=1234, fill=NA, show.legend=FALSE, size = 5) + -->
<!--             # nudge_y = c(-.1, .06, 0.06, 0.08, -0.1), -->
<!--             # nudge_x = c(-5, 0, 0, -7, 0)) + -->
<!--   labs(x = "Height in cm", y = "Median ace% per match", title = "WTA") + -->
<!--   my_theme # + -->
<!-- #   scale_color_manual("Tournament", -->
<!-- #                                 values = tournament_colors, -->
<!-- #                                 labels = c("Australian Open\n(Plexicushion)\n", -->
<!-- #                                            "French Open\n(Clay)\n", -->
<!-- #                                            "US Open\n(DecoTurf)\n", -->
<!-- #                                            "Wimbledon\n(Grass)\n")) + -->
<!-- # # + -->
<!-- # #   guides(color = guide_legend(override.aes = list(size = 7, shape=15))) +  -->
<!-- # facet_wrap(~tournament) + -->
<!-- # theme(legend.position="none") +  -->
<!-- #   ylim(c(-0.02, 0.2)) -->

<!-- p_wta_percent -->

<!-- ggarrange(p_atp_percent, NULL, p_wta_percent, common.legend = TRUE, legend="bottom", widths = c(1, 0.08, 1), nrow = 1)  -->

<!-- if(save_graph){ -->
<!-- ggsave("../plots/aces_median_percent_BW2.jpg", height=8, width=graphic_width_long) -->
<!-- } -->

<!-- ``` -->

#### Using ace percentage.... one dot per player - black white

atp_ace_ht <-   gs_players %>% 
  filter(league == "ATP") %>%
  dplyr::group_by(name) %>%
  dplyr::summarise(ht = maxish(ht),
                   avg_ave_perc = mean(ace/svpt, na.rm=TRUE),
                   med_ace_perc = median(ace/svpt, na.rm=TRUE),
                   count_n = n())

wta_ace_ht <-   gs_players %>% 
  filter(league == "WTA") %>%
  dplyr::group_by(name) %>%
  dplyr::summarise(ht = maxish(ht),
                   avg_ave_perc = mean(ace/svpt, na.rm=TRUE),
                   med_ace_perc = median(ace/svpt, na.rm=TRUE),
                   count_n = n())

p_atp_percent <- atp_ace_ht %>% # could also try median
ggplot( aes(ht, med_ace_perc)) + 
  geom_jitter(size=1.5) + 
  geom_smooth( method = "lm", color="black") + 
  geom_label_repel(data=dplyr::filter(atp_ace_ht, name %in% atp_servers),
            aes(ht, med_ace_perc, label=name),
            alpha=.8, seed=1234, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) +
  geom_label_repel(data=dplyr::filter(atp_ace_ht, name %in% atp_servers),
            aes(ht, med_ace_perc, label = name),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) +
  labs(x = "Height in cm", y = "Median ace% per match", title = "ATP") +
  my_theme 
# + 
#   ylim(c(0, 0.25))
p_atp_percent

p_wta_percent <- wta_ace_ht %>% # could also try median
ggplot( aes(ht, med_ace_perc)) + 
  geom_jitter(size=1.5) + 
  geom_smooth( method = "lm", color="black") + 
  geom_label_repel(data=dplyr::filter(wta_ace_ht, name %in% wta_servers),
            aes(ht, med_ace_perc, label=name),
            alpha=.8, seed=1234, show.legend=FALSE, size = 5,
            nudge_y = c(-.02, .01, 0.01, -0.01, -0.02 ),
            nudge_x = c(-1, 5, 5, 0, 0)) + 
  geom_label_repel(data=dplyr::filter(wta_ace_ht, name %in% wta_servers),
            aes(ht, med_ace_perc, label = name),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = c(-.02, .01, 0.01, -0.01, -0.02 ),
            nudge_x = c(-1, 5, 5, 0, 0)) +
  labs(x = "Height in cm", y = "Median ace% per match", title = "WTA") +
  my_theme
# + 
#   ylim(c(0, 0.25))

p_wta_percent

ggarrange(p_atp_percent, NULL, p_wta_percent, common.legend = TRUE, legend="bottom", widths = c(1, 0.08, 1), nrow = 1)

if(save_graph){
ggsave("../plots/aces_median_percent_BW2.jpg", height=8, width=graphic_width_long)
}

p_atp_percent <- atp_ace_ht %>% # could also try median
ggplot( aes(ht, med_ace_perc)) + 
  geom_jitter(size=1.5) + 
  geom_smooth( method = "lm", color="black") + 
  geom_label_repel(data=dplyr::filter(atp_ace_ht, name %in% atp_servers),
            aes(ht, med_ace_perc, label=name),
            alpha=.8, seed=1234, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) +
  geom_label_repel(data=dplyr::filter(atp_ace_ht, name %in% atp_servers),
            aes(ht, med_ace_perc, label = name),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = c(-.075, 0, 0, -0.05,0.02 ),
            nudge_x = c(5, 5, -5, -3, 0)) +
  labs(x = "Height in cm", y = "Median ace% per match", title = "ATP") +
  my_theme +
  ylim(c(0, 0.25))

p_wta_percent <- wta_ace_ht %>% # could also try median
ggplot( aes(ht, med_ace_perc)) + 
  geom_jitter(size=1.5) + 
  geom_smooth( method = "lm", color="black") + 
  geom_label_repel(data=dplyr::filter(wta_ace_ht, name %in% wta_servers),
            aes(ht, med_ace_perc, label=name),
            alpha=.8, seed=1234, show.legend=FALSE, size = 5,
            nudge_y = c(-.02, .01, 0.01, -0.01, -0.02 ),
            nudge_x = c(-1, 5, 5, 0, 0)) + 
  geom_label_repel(data=dplyr::filter(wta_ace_ht, name %in% wta_servers),
            aes(ht, med_ace_perc, label = name),
            seed=1234, fill=NA, show.legend=FALSE, size = 5,
            nudge_y = c(-.02, .01, 0.01, -0.01, -0.02 ),
            nudge_x = c(-1, 5, 5, 0, 0)) +
  labs(x = "Height in cm", y = " ", title = "WTA") +
  my_theme +
  ylim(c(0, 0.25))

ggarrange(p_atp_percent, p_wta_percent, common.legend = TRUE, legend="bottom",
          nrow=1)

if(save_graph){
ggsave("../plots/aces_median_percent_BW1.jpg", height=8, width=graphic_width_long)
}
